Formalizing Stack Safety as a Security Property\ifcameraready\else\vspace{-0.75cm}\fi

### 2023-02-04 submission5/core.tex
Nits

Stack safety properties need to describe the behavior of machine code. but they naturally
talk about function activations and stack contents---abstractions that
are typically not visible at machine level. To bridge this gap,
our properties are defined in terms of a {\em security semantics} layered on top of
the standard execution semantics of the machine.  The security semantics identifies certain
state transitions of the machine as {\em security-relevant
  operations}, which update
a notional {\em security context}.  This context consists of
an (abstract) stack of function activations, each associated with a {\em view}
that maps each element of the machine state (memory location or register)
to a {\em security class} (active, sealed, etc.) specifying how the activation
can access the element.
The action of a security-relevant operation on the context is defined by a set
of rules that describe how the machine code makes use of the stack and registers
to implement the function abstraction.

@@

\paragraph*{Machine model}
We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
and stack pointer. 
%, and a standard ABI.
%\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
%  This assumption is made in our example and testing.}
We make no particular assumptions about the provenance of the machine code; in particular,
we do not assume the use of any particular compiler.
The machine may possibly be enhanced with % security
enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
we assume that the behavior of these mechanisms is incorporated into the basic
step semantics of the machine, but we keep the enforcement state unchanged when constructing variant states.
Failstop behavior by enforcement mechanisms is modeled as stepping to the same state
(and thus silently diverging.)
%% \sna{I think that the payload concept is too nitty-gritty for this section.
%%   The chief use of the payload is that it allows every value to be interpreted
%%   as an address. In the tag world, it happens to coincide with a natural instance
%%   of our variant-compatibility relation.}

@@

Exactly which operations and rules are needed depends on
what code features we wish to support.
The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. A core set of operations covering calls, returns, and local
memory is introduced in the example in \cref{sec:example}
and formalized in \cref{sec:formal}. An extended set covering simple memory sharing and
tail-call elimination are described in \cref{sec:extensions} and tested in \cref{sec:testing}.
The remaining operations are needed for the capability-based model in \cref{app:ptr}.

@@

\paragraph*{Views and security classes}

@@

Conceptually, the following steps are needed to instantiate the framework to a specific machine
and code style: (i) define the base machine semantics, including any hardware
security enforcement features; (ii) identify the set of
security-relevant operations and rules required by the code style; (iii) determine
how to label machine instruction with security-relevant
operations as appropriate; (iv) specify the form of observable events.

>>>>

Stack safety properties need to describe the behavior of machine code, but they naturally
talk about function activations and stack contents---abstractions that
are typically not visible at machine level. To bridge this gap,
our properties are defined in terms of a {\em security semantics} layered on top of
the standard execution semantics of the machine.  The security semantics identifies certain
state transitions of the machine as {\em security-relevant
  operations}, which update
a notional {\em security context}.  This context consists of
an (abstract) stack of function activations, each associated with a {\em view}
that maps each element of the machine state (memory location or register)
to a {\em security class} (active, sealed, etc.) specifying how the activation
can access the element.
The action of a security-relevant operation on the context is defined by a set
of rules that describe how the machine code makes use of the stack and registers
to implement the function abstraction.

@@

\paragraph*{Machine Model}
We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
and stack pointer. 
%, and a standard ABI.
%\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
%  This assumption is made in our example and testing.}
We make no particular assumptions about the provenance of the machine code; in particular,
we do not assume the use of any particular compiler.
The machine may possibly be enhanced with % security
enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
we assume that the behavior of these mechanisms is incorporated into the basic
step semantics of the machine, but we keep the enforcement state unchanged when constructing variant states.
Failstop behavior by enforcement mechanisms is modeled as stepping to the same state
(and thus silently diverging.)
%% \sna{I think that the payload concept is too nitty-gritty for this section.
%%   The chief use of the payload is that it allows every value to be interpreted
%%   as an address. In the tag world, it happens to coincide with a natural instance
%%   of our variant-compatibility relation.}

@@

Exactly which operations and rules are needed depends on
what code features we wish to support.
The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. A core set of operations covering calls, returns, and local
memory is introduced in the example in \cref{sec:example}
and formalized in \cref{sec:formal}. An extended set covering simple memory sharing and
tail-call elimination is described in \cref{sec:extensions} and tested in \cref{sec:testing}.
The remaining operations are needed for the capability-based model in \cref{app:ptr}.

@@

\paragraph*{Views and Security Classes}

@@

Conceptually, the following steps are needed to instantiate the framework to a specific machine
and code style: (i) define the base machine semantics, including any hardware
security enforcement features; (ii) identify the set of
security-relevant operations and rules required by the code style; (iii) determine
how to label machine instructions with security-relevant
operations as appropriate; (iv) specify the form of observable events.
